Determinism, Complexity, and Predictability in Computer Performance

### 2014-04-05 results.tex
rewrite of results.tex

The experiments below involve three different prediction methods
applied to time-series data from eight different systems: {\tt
  col\_major}, {\tt 403.gcc}, and the six different regimes of the
{\tt dgesdd} signal in Figure~\ref{fig:svd-ts-colored}.  The objective of these
experiments was to explore how prediction accuracy is related to WPE,
and how that relationship depends on the generating process and the
prediction method.  Working from the first 90\% of each signal, we
generated a prediction of the last 10\% using the \naive, ARIMA, and
LMA prediction methods, as described in Section~\ref{sec:accuracy},
then calculated the MASE value of those predictions.  We also
calculated the WPE of each time series using a wordlength of six, as
described in Section~\ref{sec:meaComplex}.  In order to assess the
run-to-run variability of these results, we repeated all of these
calculations on 15 separate trials: i.e., 15 different runs of each
program.  Figure~\ref{fig:wpe_vs_mase_best} shows the {\sl best} of
these 45 predictions for each system: i.e., the lowest error over all
15 trials and all three methods for each of the eight programs.  The
WPE is plotted against the corresponding MASE value here in order to
bring out the correlation between these two quantities.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\columnwidth]{figs/prediction_vs_entropy}
  \caption{Weighted permutation entropy vs. mean absolute scaled error
    (MASE) of the best prediction of each system.
% 
% (across all trials and all prediction methods) 
% 
The dashed line is a least-squares linear fit of all the points except
{\tt dgesdd$_1$}, which we have excluded for reasons explained in the
text.}
  \label{fig:wpe_vs_mase_best}
\end{figure}

The best-case prediction error for each of these eight systems is
roughly proportional to the weighted permutation entropy, which is
consistent with our first conjecture.  The dashed line in the figure,
a linear least-squares fit to these best-case errors\footnote{with the
  exception of {\tt dgesdd$_1$}, for reasons described below} captures
this rough proportionality.  This is not a formal result.  The three
methods used here were chosen to span the space of standard prediction
strategies, but they do not cover that space exhaustively.  Our goal
here is an empirical assessment of predictability and complexity, not
formal results about a ``best'' predictor for a given time series.
There may be other methods that produce lower MASE values than those
in Figure~\ref{fig:wpe_vs_mase_best}, but the sparseness of the points
in the upper-left and lower-right quadrants of this plot strongly
suggests that the underlying predictability of a time series is
inversely correlated with its WPE.  The rest of this section describes
our results in more detail, including the measures taken to assure
meaningful comparisons across methods, trials, and programs, and
elaborates on the meaning of the dashed line in the figure.

Figure~\ref{fig:wpe_vs_mase_all} shows WPE vs. MASE plots for all 360
experiments.
\begin{figure}
  \centering
%    \includegraphics[width=\columnwidth]{figs/ARIMA_prediction_vs_entropy}
%    \includegraphics[width=\columnwidth]{figs/ARIMA_prediction_vs_entropy}
%    \includegraphics[width=\columnwidth]{figs/LMA_prediction_vs_entropy}
  \includegraphics[width=\columnwidth]{figs/predictions_vs_entropy}
\caption{WPE vs. MASE for all trials, methods, and systems. \svdone, \svdthree, and \svdfive are omitted from the top plot for scale reasons; their MASE scores are $2.676 \pm 4.328$, $31.386 \pm 0.282$, and $20.870 \pm 0.192$ respectively. The dashed line is the same as in the previous figure. Numerical values, including means and standard deviations of the errors, can be found in Table~\ref{tab:error} in the appendix.
% 
%  [[All traces except those from \svd$_1$ lie above the line,
%      indicating that LMA is better suited prediction method for the
%      computer traces considered.]]  [[This comes {\bf long} before the
%      associated discussion in the text, and it's not what this figure
%      is trying to show here.]]
% 
}
    \label{fig:wpe_vs_mase_all}
\end{figure} 
There are 15 points in each cluster, one for each trial.  (The points
in Figure~\ref{fig:wpe_vs_mase_best} are the leftmost of the points
for the corresponding system in any of these three plots.)  The WPE
values do not vary very much across trials.  For most traces, the
variance in MASE scores is low as well, resulting in small, tight
clusters.  In some cases---ARIMA predictions of {\tt col\_major}, for
instance---the MASE variance is larger, which spreads out the clusters
horizontally.  The distribution of the cluster {\sl positions} is
quite different in the three plots: clusters of predictions generated
with the nonlinear LMA method are closer to the dashed line, whereas
the naive and ARIMA prediction clusters are more spread out {\color{red}[[need to
    revisit this wording when get the third image]]}.  This is a
geometric validation of the second conjecture in this paper, as
described in the following paragraphs.

Though \col is a very simple program, its dynamics are actually quite
complicated, as discussed in Section~\ref{sec:intro}.  Recall from the
graphical representations in the right-hand column of
Figure~\ref{fig:gcc_vs_col} that the \naive and ARIMA prediction
methods do not perform as well on this signal as the nonlinear LMA
method.  The MASE values of the \naive and ARIMA predictions are $0.57
\pm$ 0.001 and $0.59 \pm 0.21$, respectively, across all 15 trials.
That is, the errors produced by these methods are a little over half
as large as the errors produced by the random-walk method---a
primitive strategy that simply uses the current value as the
prediction.  However, the WPE value for the \col trials is $0.51 \pm
0.003$, which is in the center of the complexity spectrum described in
Section~\ref{sec:intro}.  {\color{red}[[This suggests that roughly half of the
    signal is predictive structure.]]  [[I'm still uncomfortable with
    that phrasing; let's revisit it after we have the terminology and
    definitions hammered out.]]  }

This disparity---WPE values that suggest a high rate of forward
information transfer in the signal, but predictions with comparatively
poor MASE scores---is obvious in the geometry of the top two images in
Figure~\ref{fig:wpe_vs_mase_all}, where the \col clusters are far to
the right of the dashed line.  {\sl This indicates that these methods
  are not leveraging the available information in the signal.}  The
dynamics of \col may be complicated, but they are not unstructured.
This signal is nonlinear and deterministic, and if one uses a
prediction technique that is based a nonlinear model (LMA)---rather
than a method that simply predicts the running mean (\naive) or one
that uses a linear model (ARIMA)---the MASE score is much better:
$0.050 \pm 0.001$.  This prediction is 20 times more accurate than a
random-walk forecast, which is more in line with the level of
predictive structure that the low WPE value suggests is present in the
signal.  The dashed line is a heuristic that captures this argument.
It can allow practitioners to recognize when a prediction method is
not well matched to the task at hand: that is, when the time series
has more predictive structure than the method is able to use.  When
the predictions are well below that line (viz., the top images in
Figure~\ref{fig:wpe_vs_mase_all}), one should try a different method.

Note that the \col points are clustered in LMA but spread out
horizontally in ARIMA, far to the right of the dashed line (i.e., much
worse than the other methods).  {\color{red}[[We need to come up with a hypothesis
    about this spread and skew, even if it's totally tentative.  We
    may want to merge this discussion into one of the other \col
    paragraphs---or perhaps put it with the \svdone discussion later
    on?]]}

The WPE of \svdfive ($0.677 \pm 0.006$) is higher than that of {\tt
  col\_major}.  This indicates that the rate of forward information
transfer of the underlying process is lower, but that time-series data
from this system still contain a significant amount of structure that
can, in theory, be leveraged to predict the future course of the time
series.  As mentioned above, though, in the discussion of {\tt
  col\_major}, the effectiveness of any prediction strategy will
depend on how well it leverages the available information; methods
that use linear models, for instance, may fail when applied to
nonlinear processes.
% 
% (This is the basis for the second conjecture above.)  
% 
The match between model and task is an issue in the \svdfive
experiments as well.  The MASE scores of the \naive and ARIMA
predictions for this system were $20.870 \pm 0.192$ and $2.370 \pm
0.051$, respectively: that is, 20.87 and 2.37 times worse than a
simple random walk forecast of the same signals\footnote{The \naive
  MASE score is large because the amount of variance in this signal is
  high, which makes guessing the mean a particularly bad strategy.
  The same thing is true of the \svdthree signal.{\color{red}[[Joshua, it isn't the variance it is the strength in the bimodal distribution and only predicting in one mode of the distribution...]]}}.  Again, the
positions of these points on a WPE vs. MASE plot---significantly below
and to the right of the dashed line---should suggest to a practitioner
that there is more structure in this signal than the corresponding
method is able to leverage, and that a different method might do
better.  Indeed, for \svdfive, the LMA method produces a MASE score of
$ 0.718\pm 0.048 $ and a cluster of results that is much closer to the
dashed line on the WPE-MASE plot.  Again, this validates our second
conjecture: the LMA method can capture and reproduce the way in which
the \svdfive system processes information, but the \naive and ARIMA
prediction methods cannot.

The WPE of \gcc is much higher: $0.940 \pm 0.001$.  This system
transmits very little information forward in time and provides almost
no structure for prediction methods to work with.  {\color{red}[[What we see:
    \naive does the best, then LMA, then ARIMA.  Hypothesis: that that
    5\% of the structure is nonlinear, so LMA can use it better than
    ARIMA can.  \naive does better than LMA because it's filtering out
    the noise.]]}

\svdone behaves very differently than the other seven systems in this
study: though its WPE is very high ($0.95 \pm 0.02$), two of the three
prediction methods do quite well, yielding MASE scores of $0.82 \pm
0.08$ (LMA) and $0.714 \pm 0.07$ (ARIMA).  This pushes the
corresponding clusters of points in the bottom two images in
Figure~\ref{fig:wpe_vs_mase_all} well above the trend followed by the
other seven clusters.  Also, the MASE scores of the predictions of
this system produced by the \naive method are highly inconsistent:
$2.67 \pm 4.33$.  These effects, we believe, are artifacts of the way
MASE is calculated.  Recall that MASE scores are scaled relative to a
random-walk forecast, which uses the current value as the prediction.
This strategy works very badly on signals with frequent, large, rapid
transitions.  Consider a signal that oscillates from one end of its
range to the other at every step.  A signal like this will have a low
WPE, much like {\tt col\_major}.  However, a random-walk forecast of
this signal will be 180 degrees out of phase with the true
continuation.  Since random-walk error appears in the denominator of
the MASE score, this effect can shift points leftwards on a WPE
vs. MASE plot, and that is exactly why the \svdone clusters for ARIMA
and LMA are above the dashed line in Figure~\ref{fig:wpe_vs_mase_all}.
This time series, which is shown in closeup in
Figure~\ref{fig:svdone-ts},
\begin{figure}[htbp]
  \centering
    \includegraphics[width=\columnwidth]{figs/svdonets2}
\caption{A small portion of the \svdone time series}\label{fig:svdone-ts}
\end{figure} 
is not quite to the level of the worst-case signal described above,
but it still poses a serious challenge to random-walk prediction.  It
is dominated by a white-noise regime (between $\approx$1.86 and
$\approx$1.88 on the vertical scale in the Figure), punctuated by
short excursions above 1.9.  In the former regime, which makes up more
than 80\% of the signal, there are frequent dips to 1.82 and
occasional larger dips below 1.8.  These single-point dips are the
bane of random-walk forecasting.  In this particular case, roughly
40\% of the forecasted points are off by the width of the associated
dip, which skews the associated MASE scores.  Signals like this are
also problematic for the \naive prediction strategy, since the
outliers have significant influence on the mean.  This compounds the
effect of the skew in the scaling factor and creates a large spread in
the \svdone MASE values.  For all of these reasons, we view \svdone as
an outlier and exclude it from the fit calculation of the dashed line
in Figure~\ref{fig:wpe_vs_mase_best}.

@@

%is in large part one of the major findings of this work. More specifically, say we tried to predict an arbitrary noisy real-valued time series with an ``out-of-the-box" prediction strategy like ARIMA as proposed in \cite{autoArima} and say we got inconsistent and bad forecasts, (i.e., perform worse than the na\"ive random walk strategy (MASE$>1$). How do we determine if the prediction strategy is not adequate for the prediction task, or if the signal is simply too complex to predict. If a signal is too complex and too little forward information transfer is present we may not be able to do better than the random walk, in which case we should not worry ourselves over finding a more complicated prediction strategy. However, if we measure the complexity to be low, $\textrm{WPE}<0.85$ (see Fig.~\ref{fig:pred_vs_wpe}) we can most likely do much better than the random walk and should search for more adequate prediction strategies.

>>>>

The experiments below involve three different prediction methods
applied to time-series data from eight different systems: {\tt
  col\_major}, {\tt 403.gcc}, and the six different regimes of the
{\tt dgesdd} signal in Figure~\ref{fig:svd-ts-colored}.  The objective of these
experiments was to explore how prediction accuracy is related to WPE,
and how that relationship depends on the generating process and the
prediction method.  Working from the first 90\% of each signal, we
generated a prediction of the last 10\% using the \naive, ARIMA, and
LMA prediction methods, as described in Section~\ref{sec:accuracy},
then calculated the $MASE$ value of those predictions.  We also
calculated the WPE of each time series using a wordlength of six, as
described in Section~\ref{sec:meaComplex}.  In order to assess the
run-to-run variability of these results, we repeated all of these
calculations on 15 separate trials: i.e., 15 different runs of each
program.

Figure~\ref{fig:wpe_vs_mase_best} shows the {\sl best} of these 45
predictions for each system: that is, the lowest error over all 15
trials and all three methods for each of the eight programs.  The WPE
is plotted against the corresponding $MASE$ value here in order to bring
out the correlation between these two quantities.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\columnwidth]{figs/prediction_vs_entropy}
  \caption{Weighted permutation entropy vs. mean absolute scaled error
    ($MASE$) of the best prediction of each system.
% 
% (across all trials and all prediction methods) 
% 
The dashed line is a least-squares linear fit of all the points except
{\tt dgesdd$_1$}, which we have excluded for reasons explained in the
text.}
  \label{fig:wpe_vs_mase_best}
\end{figure}
The best-case prediction error for each of these eight systems is
roughly proportional to the weighted permutation entropy, which is
consistent with our first conjecture.  The dashed line in the figure,
a linear least-squares fit to these best-case errors\footnote{with the
  exception of {\tt dgesdd$_1$}, for reasons described below} captures
this rough proportionality.  This is not a formal result.  The three
methods used here were chosen to span the space of standard prediction
strategies, but they do not cover that space exhaustively.  Our goal
here is an empirical assessment of predictability and complexity, not
formal results about a ``best'' predictor for a given time series.
There may be other methods that produce lower $MASE$ values than those
in Figure~\ref{fig:wpe_vs_mase_best}, but the sparseness of the points
in the upper-left and lower-right quadrants of this plot strongly
suggests that the underlying predictability of a time series is
inversely correlated with its WPE.  The rest of this section describes
our results in more detail, including the measures taken to assure
meaningful comparisons across methods, trials, and programs, and
elaborates on the meaning of the dashed line in the figure.

Figure~\ref{fig:wpe_vs_mase_all} shows WPE vs. $MASE$ plots for all 360
experiments.
\begin{figure}
  \centering
%    \includegraphics[width=\columnwidth]{figs/ARIMA_prediction_vs_entropy}
%    \includegraphics[width=\columnwidth]{figs/ARIMA_prediction_vs_entropy}
%    \includegraphics[width=\columnwidth]{figs/LMA_prediction_vs_entropy}
  \includegraphics[width=\columnwidth]{figs/predictions_vs_entropy}
\caption{WPE vs. $MASE$ for all trials, methods, and systems. \svdone,
  \svdthree, and \svdfive are omitted from the top plot for scale
  reasons. {\color{red} All three plots need to have the same
    horizontal scale!}
% 
% ; their $MASE$ scores are $2.676 \pm 4.328$, $31.39 \pm 0.282$, and
%   $20.87 \pm 0.192$ respectively. 
% 
Numerical values, including means and standard deviations of the
errors, can be found in Table~\ref{tab:error}.  The dashed line is the
same as in the previous figure.
% 
%  [[All traces except those from \svd$_1$ lie above the line,
%      indicating that LMA is better suited prediction method for the
%      computer traces considered.]]  [[This comes {\bf long} before the
%      associated discussion in the text, and it's not what this figure
%      is trying to show here.]]
% 
}
    \label{fig:wpe_vs_mase_all}
\end{figure} 
There are 15 points in each cluster, one for each trial.  (The points
in Figure~\ref{fig:wpe_vs_mase_best} are the leftmost of the points
for the corresponding system in any of the three plots in
Figure~\ref{fig:wpe_vs_mase_all}.)  The WPE values do not vary very
much across trials.  For most traces, the variance in $MASE$ scores is
low as well, resulting in small, tight clusters.  In some
cases---ARIMA predictions of {\tt col\_major}, for instance---the
$MASE$ variance is larger, which spreads out the clusters
horizontally.  The distribution of the cluster {\sl positions} is
quite different in the three plots.  The mean $MASE$ scores of
predictions generated with the nonlinear LMA method are generally
closer to the dashed line; the ARIMA method clusters are more
widely spread, the \naive ~clusters even more so.  This geometric
validation of the second conjecture in this paper is discussed in more
details in the following paragraphs.

Though \col is a very simple program, its dynamics are actually quite
complicated, as discussed in Section~\ref{sec:intro}.  Recall from the
graphical representations in the left-hand column of
Figure~\ref{fig:forecast-example} that the \naive ~and ARIMA
prediction methods do not perform as well on this signal as the
nonlinear LMA method.  The $MASE$ scores of the \naive ~and ARIMA
predictions are $0.571 \pm 0.002$ and $0.599 \pm 0.211$, respectively,
across all 15 trials.  That is, these methods are performing only
$\approx 1.7$ times better than the random-walk method, a primitive
strategy that simply uses the current value as the prediction.
However, the WPE value for the \col trials is $0.513 \pm 0.003$, which
is in the center of the complexity spectrum described in
Section~\ref{sec:intro}.

This disparity---WPE values that suggest a high rate of forward
information transfer in the signal, but predictions with comparatively
poor $MASE$ scores---is obvious in the geometry of the top two images
in Figure~\ref{fig:wpe_vs_mase_all}, where the \col clusters are far
to the right of the dashed line.  {\sl This indicates that these
  methods are not leveraging the available information in the signal.}
The dynamics of \col may be complicated, but they are not
unstructured.  This signal is nonlinear and deterministic, and if one
uses a prediction technique that is based a nonlinear model
(LMA)---rather than a method that simply predicts the running mean
(\naive) or one that uses a linear model (ARIMA)---the $MASE$ score is
much better: $0.050 \pm 0.001$.  This prediction is 20 times more
accurate than a random-walk forecast, which is more in line with the
level of predictive structure that the low WPE value suggests is
present in the signal.

The dashed line in the Figures is a heuristic that captures the
argument at the end of the previous paragraph.  It can allow
practitioners to recognize when a prediction method is not well
matched to the task at hand: that is, when the time series has more
predictive structure than the method is able to use.  When the
predictions are well below that line (as in the top and middle images
in Figure~\ref{fig:wpe_vs_mase_all}), one should try a different
method.

Note that the \col points (blue {\color{blue}$+$} icons) are clustered
very tightly in the lower left quadrant of the LMA and \naive ~$MASE$
vs. WPE plots, but spread out horizontally in the ARIMA plot.
{\color{red}[[We need to come up with a hypothesis about this spread,
      even if it's totally tentative.  Was it that the random-walk got
      really good in a few of the \col trials?  Or that ARIMA got
      really bad in those trials?  In either case, why did that
      happen---what was it about those trials that made ARIMA bark
      and/or random walk do well?]]} The position of the LMA
cluster---well to the left of the other two, near the dashed
line---reflects this method's ability to capture and exploit the
structure that is present in this signal.

The WPE of \svdfive ($0.677 \pm 0.006$) is higher than that of {\tt
  col\_major}.  This indicates that the rate of forward information
transfer of the underlying process is lower, but that time-series data
from this system still contain a significant amount of structure that
can, in theory, be leveraged to predict the future course of the time
series.  As mentioned above, though, the effectiveness of any
prediction strategy depends on how well it leverages the available
information: methods that use linear models, for instance, may fail
when applied to nonlinear processes.
% 
% (This is the basis for the second conjecture above.)  
% 
This match between model and task is an issue in the \svdfive
experiments as well.  The $MASE$ scores of the \naive ~and ARIMA
predictions for this system are $20.870 \pm 0.192$ and $2.370 \pm
0.051$, respectively: that is, 20.87 and 2.37 times worse than a
simple random walk forecast of the same signals\footnote{The \naive
  ~$MASE$ score is large because of the bimodal nature of the
  distribution of the values of the signal, which makes guessing the
  mean a particularly bad strategy.  The same thing is true of the
  \svdthree signal.}.  As before, the positions of these points on a
WPE vs. $MASE$ plot---significantly below and to the right of the
dashed line---should suggest to a practitioner that there is more
structure in this signal than the corresponding method is able to
leverage, and that a different method might do better.  Indeed, for
\svdfive, the LMA method produces a $MASE$ score of $ 0.718\pm 0.048 $
and a cluster of results that is much closer to the dashed line on the
WPE-$MASE$ plot.  Again, this validates our second conjecture: the LMA
method can capture and reproduce the way in which the \svdfive system
processes information, but the \naive ~and ARIMA prediction methods
cannot.

The WPE of \gcc is higher still: $0.943 \pm 0.001$.  This system
transmits very little information forward in time and provides almost
no structure for prediction methods to work with.  Here, the \naive
~method performs best ($MASE=0.951 \pm 0.001$), simply because its
calculation of the mean filters out the noise in the signal.  Since
fitting a hyperplane using least squares should have similar effects,
the fact that LMA outperforms ARIMA ($1.530 \pm 0.021$ vs. $1.837 \pm
0.016$) is somewhat counterintuitive.  However, the small amount of
predictive structure that is present in this signal is nonlinear
(cf.,~\cite{mytkowicz09}), and LMA is designed to capture and exploit
that kind of structure.

\svdone behaves very differently than the other seven systems in this
study.  Though its weighted permutation entropy is very high ($0.957
\pm 0.016$), two of the three prediction methods do quite well,
yielding $MASE$ scores of $0.827 \pm 0.076$ (LMA) and $0.714 \pm
0.075$ (ARIMA).  This pushes the corresponding clusters of points in
the bottom two images in Figure~\ref{fig:wpe_vs_mase_all} well above
the trend followed by the other seven clusters.  The $MASE$ scores of
the predictions of this system produced by the \naive ~method,
however, are highly inconsistent: $2.676 \pm 4.328$.  The majority of
the blue diamond-shaped points on the top plot in the figure are
clustered near a $MASE$ score of 0.6, but three of the 15 \svdone
trials posed challenges to this prediction method, resulting in $MASE$
scores $\approx 2.5$.

These effects, we believe, are artifacts of the way $MASE$ is
calculated, as well as the nonstationarity properties of the \svdone
signal.  Recall that $MASE$ scores are scaled \emph{relative to a
  random-walk forecast}, which uses the current value as the
prediction.  This strategy works very badly on signals with frequent,
large, rapid transitions.  Consider a signal that oscillates from one
end of its range to the other at every step.  A signal like this will
have a low WPE, much like {\tt col\_major}.  However, a random-walk
forecast of this signal will be 180 degrees out of phase with the true
continuation.  Since random-walk error appears in the denominator of
the $MASE$ score, this effect can shift points leftwards on a WPE
vs. $MASE$ plot, and that is exactly why the \svdone clusters for
ARIMA and LMA are above the dashed line in
Figure~\ref{fig:wpe_vs_mase_all}.  This time series, which is shown in
closeup in Figure~\ref{fig:svdone-ts},
\begin{figure}[htbp]
  \centering
    \includegraphics[width=\columnwidth]{figs/svdonets2}
\caption{A small portion of the \svdone time series}\label{fig:svdone-ts}
\end{figure} 
is not quite to the level of the worst-case signal described above,
but it still poses a serious challenge to random-walk prediction.  It
is dominated by a white-noise regime (between $\approx$1.86 and
$\approx$1.88 on the vertical scale in the Figure), punctuated by
short excursions above 1.9.  In the former regime, which makes up more
than 80\% of the signal, there are frequent dips to 1.82 and
occasional larger dips below 1.8.  These single-point dips are the
bane of random-walk forecasting.  In this particular case, roughly
40\% of the forecasted points are off by the width of the associated
dip, which skews the associated $MASE$ scores.  Signals like this are
also problematic for the \naive ~prediction strategy, since the
outliers have significant influence on the mean.  This compounds the
effect of the skew in the scaling factor and creates a large spread in
the \svdone $MASE$ values.  {\color{red} We need to think about the
  three outliers and put in a sentence here about them.}  For all of
these reasons, we view \svdone as an outlier and exclude it from the
fit calculation of the dashed line in
Figure~\ref{fig:wpe_vs_mase_best}.

@@

%is in large part one of the major findings of this work. More specifically, say we tried to predict an arbitrary noisy real-valued time series with an ``out-of-the-box" prediction strategy like ARIMA as proposed in \cite{autoArima} and say we got inconsistent and bad forecasts, (i.e., perform worse than the na\"ive random walk strategy ($MAS>1$). How do we determine if the prediction strategy is not adequate for the prediction task, or if the signal is simply too complex to predict. If a signal is too complex and too little forward information transfer is present we may not be able to do better than the random walk, in which case we should not worry ourselves over finding a more complicated prediction strategy. However, if we measure the complexity to be low, $\textrm{WPE}<0.85$ (see Fig.~\ref{fig:pred_vs_wpe}) we can most likely do much better than the random walk and should search for more adequate prediction strategies.
